\section{Conclusion}
\label{sec:conclusion}

We analyzed five real-world networks across three datasets using a
consistent pipeline of centrality analysis, clustering and path-length
measurement, null-model comparison, and community detection with two
algorithms.
A recurring finding is the \textit{hub--bridge duality}: the nodes with the
highest degree (or eigenvector centrality) are not the same nodes that
maximize information flow across the network.
Top-10\% intersection analysis makes this concrete — In-Deg $\cap$ Out-Deg
is the smallest pairwise overlap in every Higgs network, confirming that
passive popularity and active broadcasting are distinct roles, while
Betweenness $\cap$ Eigenvector is the largest overlap in the Retweet graph
(18{,}628 nodes), indicating that bridges and recursively influential nodes
coincide in broadcast-heavy topologies.

The three datasets occupy distinct structural regimes.
The Higgs networks are sparse and broadcast-oriented: low clustering
($\bar{C} = 0.007$--$0.087$) with short average path lengths
($\ell \approx 5$--$7$), dominated by a single hub (node \texttt{88}) whose
in-degree and eigenvector centrality are an order of magnitude above the
next-ranked node.
The Enron email network is hierarchical: moderate clustering
($\bar{C} = 0.174$), 252 Louvain communities whose size distribution mirrors
the corporate org chart, and a COO (\texttt{sally.beck@enron.com}) who serves
simultaneously as the top in-degree node and a high-betweenness bridge — a
rare violation of hub--bridge duality that reflects the centralised
communication patterns of a corporate hierarchy.
The Facebook ego-network is small-world: the highest clustering of any
network studied ($\bar{C} = 0.606$) paired with the shortest path length
($\ell \approx 3.7$), and 16 Louvain communities that align with the known
ego-circle structure.

Comparison against Erd\H{o}s--R\'enyi, Watts--Strogatz, and
Barab\'asi--Albert null models confirms that no single generative mechanism
explains any of these networks.
Facebook's clustering approaches Watts--Strogatz, but its degree distribution
follows a power-law tail matching Barab\'asi--Albert.
Louvain consistently outperforms Label Propagation in modularity
($Q = 0.835$ vs.\ $0.737$ on Facebook; $0.628$ vs.\ $0.382$ on Enron),
and Label Propagation over-fragments by $1.3$--$1.8\times$, producing
partitions that are harder to interpret semantically.

Taken together, these results demonstrate that a multi-metric approach
— combining centrality, global graph properties, null-model benchmarking,
and community detection — is necessary to characterise the structural
roles and organisational logic of real-world networks.
Degree and betweenness capture complementary aspects of node importance;
clustering and path length distinguish network types; and algorithm choice
materially affects the granularity of discovered communities.

\section{Future Work}
\label{sec:future}

Several extensions of this work are natural:

\textbf{Temporal analysis.}
The Higgs dataset includes activity timestamps in
\texttt{higgs-activity\_time.txt}.
A natural follow-up is to study how centrality rankings evolve over time and
whether early adopters of the Higgs boson story maintain their influence.

\textbf{Multiplex centrality.}
Because mention, reply, and retweet networks share the same node set, one
could compute \textit{multiplex centrality} measures that aggregate
information across layers, identifying nodes that are simultaneously
important across all interaction types.

\textbf{Robustness and targeted removal.}
Given the hub--bridge distinction, a comparative simulation of random vs.\
targeted node removal (targeting hubs vs.\ bridges) would quantify the
resilience of each network and test whether the duality has practical
consequences for network robustness.

\textbf{Feature-based community interpretation.}
The Facebook dataset includes rich node feature vectors (\texttt{node\_to\_features.txt}).
Linking detected communities to demographic or interest features could
validate the semantic meaning of the Louvain partitions and explain why
Label Propagation over-fragments relative to the ego-circle ground truth.
